# User-Defined Models

The **subdiff** tools can be applied to a large family of user-defined models.   These are based on the location-scale modeling framework of @lysy.etal16, whereby the particle trajectory $\XX(t)$ is given by
\begin{equation}
  \XX(t) = \RR(t \mid \pph)' \mmu + \SSi^{1/2} \ZZ(t),
  (\#eq:locscale)
\end{equation}
where:

- $\ZZ(t) = \big(Z_1(t), \ldots, Z_{\ndim}(t)\big)$ are independent and identically distributed (iid) copies of a CSI process $Z(t)$ having MSD

	$$
	\msd_Z(t) = \eta(t \mid \pph).
	$$

- $\SSi$ is a $\ndim \times \ndim$ symmetric positive-definite scaling matrix.

- $\RR(t \mid \pph) = \big( R_1(t \mid \pph), \ldots, R_{\ndrift}(t \mid \pph) \big)$ is a set of $\ndrift$ drift basis functions.  Typically it will be linear or quadratic, corresponding to $\RR(t) = t$ and $\RR(t) = (t, t^2)$ with $\ndrift = 1$ and $\ndrift = 2$, respectively.  However, **subdiff** allows for an arbitrary number of basis functions possibly dependent on the "kernel" parameters $\pph$.

- $\mmu$ is $\ndrift \times \ndim$ matrix of drift coefficients.  For linear drift, $\mmu = (\rv \mu \ndim)$ correspond to per-coordinate drift velocities.

An attractive feature of model \@ref(eq:locscale) is that the MSD of the drift-subtracted particle trajectory $\tilde{\XX}(t) = \XX(t) - \RR(t)' \mmu$ has the simple formula

$$
\msd_{\tilde{\XX}} = \tfrac 1 {\ndim} \tr(\SSi) \cdot  \eta(t \mid \pph).
$$

Another appealing property of this model is that its parameters $\tth = (\pph, \mmu, \SSi)$ can be estimated from discrete observations $\XX = (\rv [0] {\XX} N)$ in a computationally efficient manner [@ling.etal20].

## Example: Brownian Motion in a Jeffreys Fluid

Consider a particle of radius $r$ in a Jeffreys fluid described by viscosity parameters $\eta_1$ and $\eta_2$ and elastic modulus $\Gamma$.  The MSD of the drift-subtracted particle trajectory $\tilde{\XX}(t)$ is given by [@raikher.rusakov10]

\begin{equation}
\msd_{\tilde{\XX}}(t) = \frac{2 \kbt}{\xi_V(1+q)} \times \left\{t + \frac{q}{1+q} \tau_M \left[1 - \exp\left(-\frac{(1+q)t}{\tau_M}\right)\right]\right\},
(\#eq:jfsmsd)
\end{equation}

where $q = \eta_1/\eta_2$, $\tau_M = (\eta_1 + \eta_2)/\Gamma$, and $\xi_V = 6\pi r \cdot \eta_1\eta_2/(\eta_1 + \eta_2)$.  A more natural parametrization is

$$
\msd_{\tilde{\XX}}(t) = \sigma^2 \left\{t + \alpha \left[1 - e^{-t/\tau}\right]\right\},
$$

where $\tau = \tau_M/(1+q) = \eta_2/\Gamma$, $\alpha = \tau q = \eta_1/\Gamma$, and $\sigma^2 = 2\kbt / (\xi_V(1+q)) = 2\kbt / (6\pi r \eta_1)$.  In the formulation of the location-scale model \@ref(eq:locscale), the CSI process $Z(t)$ underlying the particle trajectory $\XX(t)$ has MSD

$$
\eta(t \mid \pph) = \msd_Z(t) = \left\{t + \alpha \left[1 - e^{-t/\tau}\right]\right\}
$$

with $\pph = (\alpha, \tau)$, and the MSD scale factor is $\sigma^2 = \tr(\SSi)/\ndim$.

## Creating the Jeffreys Model Class

In order to fit the Jeffreys model \@ref(eq:jfsmsd) to trajectory data, we will create a class `jfs_model` which will derive from the base class `csi_model` used to represent the general location-scale model \@ref(eq:locscale).  In order to do this, we will need the following ingredients:

- The increment autocorrelation function

	$$
	\acf_{\Delta Z}(h) = \cov(\Delta Z_n, \Delta Z_{n+h}).
	$$

	This must be supplied as a function with arguments $(\pph, \dt, N)$ and return the vector $\big(\acf_{\Delta Z}(0), \ldots, \acf_{\Delta Z}(N-1)\big)$.  Note that $\msd_{Z}(t)$ can be converted to $\acf_{\Delta Z}(h)$ via the function `SuperGauss::msd2acf()`.

- Functions to convert back and forth between the kernel parameters in the "inferential" basis, $\pph$, and these parameters in a "computational basis", $\pps$.  That is, $\pps$ should have the same dimension as $\pph$ but have no constraints.  For the Jeffreys model, since $\alpha, \tau > 0$, a natural choice of computational basis is $\pps = (\log \alpha, \log \tau)$.

- Specification of the drift basis functions $\RR(t \mid \pph)$.  This is done automatically in the case of linear, quadratic, or no drift, so we'll skip this here.

- A character vector of the names of the elements of $\pph$ for display purposes.

The following code block defines the `jfs_model` class.  This is done with the [R6](https://r6.r-lib.org/index.html) class mechanism, of which a basic understanding is assumed.  More information about the base class for location-scale models can be obtained from the package documentation: `?csi_model`.

```{r jfs_model}
#' Calculate the ACF of the Jeffreys model.
#'
#' @param alpha Scale parameter.
#' @param tau Decorrelation time.
#' @param dt Interobservation time.
#' @param N Number of ACF lags to return.
#'
#' @return An ACF vector of length `N`.
jfs_acf <- function(alpha, tau, dt, N) {
  t <- (1:N)*dt # time vector
  msd <- (t + alpha * (1 - exp(-t/tau)))
  SuperGauss::msd2acf(msd) # convert msd to acf
}

jfs_model <- R6::R6Class(
  classname = "jfs_model",
  inherit = subdiff::csi_model,

  public = list(
    # Kernel parameter names in the inferential basis.
    phi_names = c("alpha", "tau"),

    # Autocorrelation function.
    acf = function(phi, dt, N) {
      jfs_acf(alpha = phi[1], tau = phi[2], dt = dt, N = N)
    },

    # Transform kernel parameters from inferential to computational basis.
    trans = function(phi) {
      psi <- log(phi)
      names(psi) <- private$psi_names # these are determined automatically.
      psi
    },

    # Transform kernel parameters from computational to inferential basis.
    itrans = function(psi) {
      phi <- exp(psi)
      names(phi) <- public$phi_names
      phi
    }
  )
)
```

## Simulation

```{r jfs_sim_calc, include = FALSE}
kB <- 1.38064852e-23 # Boltzmann constant

# simulation parameters
tau_M <- 2.1 # seconds
q <- 4.6e1 # unitless
xi_V <- 1.4e-07 # kg/s
rad <- 0.48 # particle radius (micron)
Temp <- 298 # temperature of the system (Kelvin)

dt <- 1/60 # interobservation time
N <- 1800 # number of observations
ndim <- 2 # number of dimensions

# increment autocorrelation
tau <- tau_M/(1+q)
alpha <- tau * q
acf <- jfs_acf(alpha = alpha, tau = tau, dt = dt, N = N)
# increment drift matrix: in this case zero drift
drift <- matrix(0, N, ndim)
# scale matrix
sigma2 <- 2*kB*Temp / (xi_V * (1+q))
Sigma <- sigma2 * diag(ndim)
# position at time t = 0
X0 <- rep(0, ndim)

# generate data
Xt <- csi_sim(drift = drift, acf = acf,
              Sigma = Sigma, X0 = X0)

# plot empirical and theoretical MSD
tmax <- 10 # maximum time to plot
nmax <- floor(tmax/dt) # corresponding MSD lag number
tibble(
  Time = 1:nmax * dt, # time sequence
  Theoretical = sigma2 * SuperGauss::acf2msd(acf[1:nmax]), # theoretical msd
  Empirical = msd_fit(Xt, nlag = nmax) # empirical msd
) %>%
  pivot_longer(cols = Theoretical:Empirical, names_to = "MSD Type",
               values_to = "MSD", values_ptypes = list(factor())) %>%
  ggplot(aes(x = Time, y = MSD * 1e12, group = `MSD Type`)) +
  geom_line(aes(color = `MSD Type`)) +
  scale_x_log10() + scale_y_log10() +
  xlab(expression("Time "*(s))) + ylab(expression("MSD "*(mu*m^2)))

```

<!-- In @raikher.rusakov10, the Jeffreys model \@ref(eq:jfsmsd) is calibrated to particles of radius $r = `r rad`\, \mathrm{\mu m}$ in a solution of cetyltrimethyl ammonium bromide $[\mathrm{CTAB}] = 1\, \mathrm{g}\, \mathrm{cm}^{-3}$ and potassium bromide $[\mathrm{KBr}] = 2\, \mathrm{M}$ at temperature $T = `r Temp`\, \mathrm{K}$.  The fitted parameters are $\tau_M = `r tau_M`\, \mathrm{s}$, $q = `r q`$, and $\xi_V = `r xi_V`\, \mathrm{kg/s}$. -->

The following code shows how to simulate the trajectory of a particle from the Jeffreys model with parameters $\tau_M = `r tau_M`\, \mathrm{s}$, $q = `r q`$, and $\xi_V = `r xi_V`\, \mathrm{kg/s}$ and temperature $T = `r Temp`\, \mathrm{K}$.  The 2D trajectory is generated with zero drift $\RR(t \mid \pph) = \bm{0}$, and scale factor $\SSi = \left[\begin{smallmatrix} \sigma^2 & 0 \\ 0 & \sigma^2 \end{smallmatrix} \right]$.  Figure \@ref(fig:jfssim) compares the empirical MSD for the simulated trajectory to the theoretical MSD.

(ref:jfssim) Empirical and theoretical MSD for the Jeffreys model with parameters $\tau_M = `r tau_M`\, \mathrm{s}$, $q = `r q`$, and $\xi_V = `r xi_V`\, \mathrm{kg/s}$ and temperature $T = `r Temp`\, \mathrm{K}$.

```{r jfssim, ref.label = "jfs_sim_calc", fig.width = 7, fig.height = 3, fig.cap = "(ref:jfssim)"}
```

## Parameter Estimation



```{r jfs_fit}
mod <- jfs_model$new(Xt = Xt, dt = dt)
```


```{r scratch, eval = FALSE, include = FALSE}
#' Convert original parameters to standardized parameters for the Jeffreys model.
#'
#' @param eta1 First viscosity parameter (positive real).
#' @param eta2 Second viscosity parameter (positive real).
#' @param Gamma The elastic modulus of the fluid (positive real).
#' @param Temp Temperature of the system.
#'
#' @return Vector of standardized parameters `(tau_M, q, xi_V)`.
jfs_stpar <- function(eta1, eta2, Gamma, Temp) {
  
}


#' Calculate the MSD of the Jeffreys model.
#'
#' @param t Vector of timepoints at which to calculate the MSD.
#' @param r Radius of the particle.
#' @param eta1 First viscosity parameter.
#' @param eta2 Second viscosity parameter.
#' @param Gamma The elastic modulus of the fluid.
#' @param Temp Temperature of the system.
#'
#' @return MSD vector of the same length as `t`.
jfs_msd <- function(t, r, eta1, eta2, Gamma, Temp) {
  Kb <- 1.381e-23 # Boltzmann's constant
  tau <- eta2/Gamma
  alpha <- eta1/Gamma
  sigma2 <- 2 * kb * Temp / (6 * pi * r * eta1)
  sigma2 * (t + alpha * (1 - exp(-abs(t)/tau)))
}

# Calculate the standardized parametrization
jfs_std <- function(alpha, tau, sigma2, Temp) {
  kB <- 1.38064852e-23 # Boltzmann constant ( m^2 kg /(s^2 K) )
  q <- alpha/tau
  tau_M <- tau * (1+q)
  xi_V <- 2*kB*Temp/(sigma2 * (1+q))
  setNames(c(tau_M, q, xi_V), c("tau_M", "q", "xi_V"))
}
# Calculate the natural parametrization
jfs_nat <- function(tau_M, q, xi_V, Temp) {
  kB <- 1.38064852e-23 # Boltzmann constant
  sigma2 <- 2*kB*Temp/(xi_V * (1+q))
  alpha <- q/(1+q) * tau_M
  tau <- tau_M/(1+q)
  setNames(c(alpha, tau, sigma2), c("alpha", "tau", "sigma2"))
}
# Calculate MSD
jfs_msd <- function(t, tau_M, q, xi_V, Temp) {
  ## kB <- 1.38064852e-23 # Boltzmann constant
  ## sigma2 <- 2*kB*Temp/(xi_V * (1+q))
  ## alpha <- q/(1+q) * tau_M
  ## tau <- tau_M/(1+q)
  theta_nat <- jfs_nat(tau_M, q, xi_V, Temp)
  alpha <- theta_nat[1]
  tau <- theta_nat[2]
  sigma2 <- theta_nat[3]
  sigma2 * (t + alpha * (1 - exp(-t/tau)))
}

tau_M <- 1.8e1 # seconds
q <- 6.7e1 # unitless
xi_V <- 1.2e-10 # kg/s
rad <- 0.48 # particle radius (micron)
Temp <- 298 # temperature of the system (Kelvin)
theta_std <- c(tau_M = tau_M, q = q, xi_V = xi_V)
theta_nat <- do.call(jfs_nat, as.list(c(theta_std, Temp = Temp)))
do.call(jfs_std, as.list(c(theta_nat, Temp = Temp))) - theta_std

tseq <- exp(seq(log(1/60), log(30), len = 1000))
alpha <- 8.2
tau <- 4.6e-2
sigma2 <- 9.1e-15
plot(tseq, 1e12 * sigma2 * (tseq + alpha * (1 - exp(-tseq/tau))),
     type = "l", log = "xy", xlab = "Time (s)", ylab = "MSD (microns^2)")

```
