---
title: "**subdiff**: Tools for the Analysis of Passive Particle-Tracking Data"
author: "Martin Lysy, Yun Ling"
date: "`r Sys.Date()`"
output: 
  bookdown::html_vignette2:
    toc: true
    toc_depth: 4
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    toc_float: true
    #code_folding: hide
    #keep_md: yes
bibliography: references.bib
csl: taylor-and-francis-harvard-x.csl
# csl: journal-of-the-american-statistical-association.csl
link-citations: true
vignette: >
  %\VignetteIndexEntry{Introduction to subdiff}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({
      TeX: {
	  Macros: {
	      rv: ["#2_{#1},\\ldots,#2_{#3}",3,"1"]
	  }
      }
  });
</script>


\newcommand{\bm}[1]{\boldsymbol{#1}}
\newcommand{\tx}[1]{\textrm{#1}}
<!-- \newcommand{\rv}[3][1]{#2_{#1},\ldots,#2_{#3}} -->
\newcommand{\RR}{\bm{R}}
\newcommand{\BB}{\bm{B}}
\newcommand{\JJ}{\bm{J}}
\newcommand{\XX}{\bm{X}}
\newcommand{\UU}{\bm{U}}
\newcommand{\VV}{\bm{V}}
\newcommand{\dX}{\Delta\XX}
\newcommand{\YY}{\bm{Y}}
\newcommand{\ZZ}{\bm{Z}}
\newcommand{\dt}{\Delta t}
\newcommand{\tmin}{t_{\tx{min}}}
\newcommand{\tmax}{t_{\tx{max}}}
\newcommand{\aD}{(\alpha,D)}
\newcommand{\eps}{\varepsilon}
\newcommand{\N}{\mathcal{N}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\msd}{\operatorname{MSD}}
\newcommand{\acf}{\operatorname{ACF}}
\newcommand{\var}{\operatorname{var}}
\newcommand{\se}{\operatorname{se}}
\newcommand{\cov}{\operatorname{cov}}
\newcommand{\tr}{\operatorname{trace}}
\newcommand{\arma}{\operatorname{ARMA}}
\newcommand{\snr}{\mathrm{SNR}}
\newcommand{\kbt}{k_{\mathrm{B}}T}
\newcommand{\argmax}{\operatorname{arg\,max}}
\newcommand{\mmu}{\bm{\mu}}
\newcommand{\eet}{\bm{\eta}}
\newcommand{\tth}{\bm{\theta}}
\newcommand{\lla}{\bm{\lambda}}
\newcommand{\oom}{\bm{\omega}}
\newcommand{\pps}{\bm{\psi}}
\newcommand{\pph}{\bm{\phi}}
\newcommand{\zze}{\bm{\zeta}}
\newcommand{\SSi}{\bm{\Sigma}}
\newcommand{\ndim}{d}
\newcommand{\ndrift}{k}
\newcommand*\ud{\mathop{}\!\mathrm{d}}

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# silently load packages not used in code
suppressMessages({
  require(kableExtra)
})
# format link to CRAN packages
cran_link <- function(pkg) {
  paste0("[**", pkg, "**](https://CRAN.R-project.org/package=", pkg, ")")
}
```

# Introduction

The **subdiff** package provides a number of utilities for analyzing subdiffusion in passive particle tracking data.  Let $\XX(t) = \big(X_1(t), \ldots, X_{\ndim}(t)\big)$ denote the $\ndim$-dimensional position of a particle at time $t$, where $\ndim = 1-3$.  In an ideal experiment, it is typically assumed that $\XX(t)$ is a Continuous Stationary Increments (CSI) process, i.e., a stochastic process with mean zero and such that the distribution of $\dX(t) = \XX(t+\dt) - \XX(t)$ is independent of $t$.  Under these conditions, the covariance properties of $\XX(t)$ can often be deduced entirely from the particle's mean square displacement (MSD),
\begin{equation}
\msd_{\XX}(t) = \frac 1 \ndim \sum_{i=1}^\ndim E\big[\big(X_i(t)-X_i(0)\big)^2\big].
(\#eq:msd)
\end{equation}
A widely observed phenomenon is that the MSD of microparticles diffusing in biological fluids has a power-law signature on a given timescale,
\begin{equation}
  \msd_{\XX}(t) \sim D \cdot t^\alpha, \qquad t \in (\tmin, \tmax).
  (\#eq:subdiff)
\end{equation}
Whereas $\alpha = 1$ corresponds to the MSD of ordinary diffusion observed in Brownian particles, the power law MSD \@ref(eq:subdiff) with $\alpha \in (0,1)$ is referred to as *subdiffusion*.  

The purpose of the **subdiff** library is to provide simple and computationally efficient tools for modeling and estimating subdiffusion.  An overview of these tools is presented below.  

## Installation

```{r load_packages, include = FALSE, message = FALSE}
# load required packages
require(subdiff)
require(tidyr)
require(dplyr)
require(optimCheck)
require(ggplot2)
## require(RColorBrewer)
require(gridExtra)
require(parallel)
```

```{r ref.label = "load_packages"}
```

```{r hbe_info, include = FALSE}
npaths <- length(unique(hbe$id)) # number of trajectories
n_rng <- range(tapply(hbe$id, hbe$id, length)) # range of trajectory lengths
dt <- 1/60 # interobservation time
```

```{r dataset, child = "_dataset.Rmd"}
```

```{r semiparam, child = "_semiparametric.Rmd"}
```

```{r param, child = "_parametric.Rmd"}
```

```{r csiclass, child = "_csiclass.Rmd"}
```

# References

