---
title: "Visual Check of Subdiffusion Estimators"
author: "Martin Lysy"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: yes
bibliography: references.bib
csl: technometrics.csl
link-citations: true
vignette: >
  %\VignetteIndexEntry{Visual Check of Subdiffusion Estimators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

\newcommand{\bm}[1]{\boldsymbol{#1}}
\newcommand{\tx}[1]{\mathrm{#1}}
\newcommand{\XX}{\bm{X}}
\newcommand{\YY}{\bm{Y}}
\newcommand{\dt}{\Delta t}
\newcommand{\N}{\mathcal{N}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\msd}{\operatorname{MSD}}
\newcommand{\var}{\operatorname{var}}

```{r setup}
require(subdiff)
require(dplyr)
require(tidyr)
require(ggplot2)
require(scales)
```

# Motivation

The purpose of this vignette is to simulate data from simple subdiffusion models and visually assess that various candidate estimators produce the correct value.

The MSD of a $d$-dimensional trajectory $\XX(t) = (X_1(t), \ldots, X_d(t))$ is defined as
$$
\msd(t) = \sum_{i=1}^d E[(X_i(t) - X_i(0))^2],
$$
and the MSD of fractional Brownian motion with parameters $(\alpha, D)$ is defined as
$$
\msd_{\tx{fBM}}(t) = 2d \cdot D t^\alpha.
$$

# Ordinary Brownian Motion

Let us verify the conventions above by estimating the parameters of simulated trajectories of Brownian motion.  The data will be generated and analysed in $d = 2$ dimensions.

First let us compare the empirical MSD estimate to the true function $\msd(t) = 4 D t$.

```{r bm_msd, fig.width = 7, fig.height = 3.5, out.width = "100%"}
ndim <- 2 # number of trajectory dimensions
# true parameter values
D <- .43
mu <- runif(ndim) # drift per coordinate
# simulation setup
N <- 1800 # number of observations
dt <- 1/60 # framerate

# simulate a trajectory
dX <- t(matrix(rnorm(N*ndim), ndim, N) * sqrt(2 * D * dt) + mu * dt)
Xt <- apply(dX, 2, cumsum)

msd_hat <- cbind(nodrift = msd_fit(Xt, nlag = nrow(Xt)-2),
                 direct = msd_fit(Xt, nlag = nrow(Xt)-2, demean = FALSE))
tibble(t = 1:nrow(msd_hat) * dt,
       direct = msd_hat[,"direct"],
       nodrift = msd_hat[,"nodrift"],
       true = 2*ndim * D * t) %>%
  pivot_longer(cols = direct:true, names_to = "type",
               values_to = "msd") %>%
  mutate(type = factor(type, levels = c("true", "direct", "nodrift"),
                       labels = c("'True '*MSD(t) == 4*D*t",
                                  "'Estimate (Direct)'",
                                  "'Estimate (Drift sub.)'"))) %>%
  ggplot(aes(x = t, y = msd)) +
  geom_line(aes(color = type)) +
  scale_x_log10() + scale_y_log10() +
  scale_color_discrete(labels = label_parse()) +
  xlab(expression("Time "*"(s)")) +
  ylab(expression("MSD "*(mu*m^2))) +
  theme(legend.position = c(0,1),
        legend.title = element_blank(),
        legend.justification = c(-.02, 1.02),
        legend.text.align = 0)
```

Now let us compare various subdiffusion estimators to the true values of $(\alpha, D)$.  The subdiffusion estimators are:

- `ls`: A version of least-squares.  There is no unique definition of the LS estimator as it depends among other things on the method of drift correction and the set of timepoints used in the regression.  Here we use a linear drift subtraction and timepoints $t = \dt \cdot (1, 2, 5, 10, 20, 50, 100, 200, 500)$.

- `fbm`: MLE of fractional Brownian motion.  Drift correction in this and the subsequent MLE estimators is linear.

<!-- - MLE of fBM + white noise floor (`fSN`). -->

- `fsd`: MLE of fBM + Savin-Doyle noise model [@savin.doyle05].  This accounts for static noise via a white noise floor and dynamic errors due to particle movement during the camera exposure time for each frame.  It's identical to fSN when there is no dynamic error (camera exposure time = 0).

- `fma`: MLE with fBM + MA(1) noise.  This is one of the models proposed in @ling.etal19.

- `farma`: MLE with fBM + ARMA(1,1) noise, another model proposed in @ling.etal19.

Since there is no noise here and the fBM model is correct -- in that BM is a special case of it with $\alpha = 1$ -- we expect all estimators to produce very similar results.  The code below shows that the **subdiff** implementation of these models consistently estimates $(\alpha, D)$.

```{r bm_fit}
#' Calculate `(alpha, log(D))` estimates for various estimators.
#'
#' @param Xt Matrix of `N x ndim` particle trajectory observations.
#' @param dt Interobservation time.
#' @return Matrix with two rows corresponding to `alpha` and `D` estimates for models: "ls", "fbm", "fsd", "fma", and "farma".
fit_models <- function(Xt, dt) {
  ls_lags <- c(1, 2, 5, 10, 20, 50, 100, 200, 500)
  ad <- cbind(
    ls = ls_fit(Xt, dt, lags = ls_lags, vcov = FALSE),
    fbm = fbm_fit(Xt, dt, vcov = FALSE),
    fsd = fsd_fit(Xt, dt, vcov = FALSE),
    fma = farma_fit(Xt, dt, order = c(0, 1), vcov = FALSE),
    farma = farma_fit(Xt, dt, order = c(1, 1), vcov = FALSE)
  )
  ad[2,] <- exp(ad[2,]) # convert logD to D
  rownames(ad)[2] <- "D"
  ad
}

ad_fit <- fit_models(Xt, dt)
# display alongside true value
signif(cbind(True = c(alpha = 1, D = D), ad_fit), 2)
```

# Fractional Brownian Motion

This just confirms that the estimators remain consistent with simulated [fractional Brownian motion](https://en.wikipedia.org/wiki/Fractional_Brownian_motion) (fBM).  First, we simulate an fBM trajectory and plot the true and empirical MSD.  Note that, for testing purposes, the fBM trajectory is generated via the Cholesky decomposition of its variance matrix.  This method is transparent but scales cubically in the number of fBM observations.  A more efficient method of simulation is used by `SuperGauss::rnormtz()`, which is also used internally by **subdiff**.

```{r fbm_msd, fig.width = 7, fig.height = 3.5, out.width = "100%"}
ndim <- 2 # number of trajectory dimensions
# true parameter values
D <- .43
alpha <- runif(1) # randomly chosen between 0 and 1
mu <- runif(ndim) # drift per coordinate
# simulation setup
N <- 1800 # number of observations
dt <- 1/60 # framerate

# simulate a trajectory
# do this using an inefficient but easy to check method
V <- outer(
  X = 1:N * dt,
  Y = 1:N * dt,
  FUN = function(t, s) {
    D * (abs(t)^alpha + abs(s)^alpha - abs(t-s)^alpha)
  }
)
Xt <- t(chol(V)) %*% matrix(rnorm(N*ndim), N, ndim)
Xt <- Xt + (1:N * dt) %o% mu
## dX <- SuperGauss::rnormtz(n = ndim,
##                           acf = 2 * D * fbm_acf(alpha = alpha, dT = dt, N = N))
## dX <- sweep(dX, 2, mu * dt, FUN = "+")
## Xt <- apply(dX, 2, cumsum)

# msd estimate
msd_hat <- cbind(nodrift = msd_fit(Xt, nlag = nrow(Xt)-2),
                 direct = msd_fit(Xt, nlag = nrow(Xt)-2, demean = FALSE))
tibble(t = 1:nrow(msd_hat) * dt,
       direct = msd_hat[,"direct"],
       nodrift = msd_hat[,"nodrift"],
       true = 2*ndim * D * t^alpha) %>%
  pivot_longer(cols = direct:true, names_to = "type",
               values_to = "msd") %>%
  mutate(type = factor(type, levels = c("true", "direct", "nodrift"),
                       labels = c("'True '*MSD(t) == 4*D*t^alpha",
                                  "'Estimate (Direct)'",
                                  "'Estimate (Drift sub.)'"))) %>%
  ggplot(aes(x = t, y = msd)) +
  geom_line(aes(color = type)) +
  scale_x_log10() + scale_y_log10() +
  scale_color_discrete(labels = label_parse()) +
  xlab(expression("Time "*"(s)")) +
  ylab(expression("MSD "*(mu*m^2))) +
  theme(legend.position = c(0,1),
        legend.title = element_blank(),
        legend.justification = c(-.02, 1.02),
        legend.text.align = 0)
```

Now we fit the fBM trajectory wit various subdiffusion estimators:

```{r fbm_fit}
# fit various estimators
ad_fit <- fit_models(Xt, dt)

# display estimates with true value
signif(cbind(True = c(alpha = alpha, D = D), ad_fit), 2)
```

```{r scratch, echo = FALSE, eval = FALSE}
log1pe <- subdiff:::log1pe
model <- fsd_model$new(Xt, dt)
penalty <- function(psi) log1pe(psi[2]) + log1pe(-psi[2]) - psi[3]
objfun <- function(psi) model$nlp(psi) + penalty(psi)

fit <- optim(par = rep(0, 3),
             fn = objfun)

fsd_fit(Xt, dt, vcov = FALSE)

optim_proj(fit$par, model$nlp, maximize = FALSE, xrng = .5)
```
