---
title: "Introduction to **subdiff**: Tools for the Analysis of Passive Particle-Tracking Data"
author: "Martin Lysy, Yun Ling"
date: "`r Sys.Date()`"
pkgdown:
  as_is: true
output: 
  bookdown::html_vignette2:
    toc: true
  bookdown::html_document2:
    toc: true
    toc_float: true
    #code_folding: hide
    #keep_md: yes
bibliography: references.bib
csl: technometrics.csl
link-citations: true
vignette: >
  %\VignetteIndexEntry{Introduction to **subdiff**: Tools for the Analysis of Passive Particle-Tracking Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- <script type="text/x-mathjax-config"> -->
<!--   MathJax.Hub.Config({ -->
<!--       TeX: { -->
<!-- 	  Macros: { -->
<!-- 	      rv: ["#2_{#1},\\ldots,#2_{#3}",3,"1"] -->
<!-- 	  } -->
<!--       } -->
<!--   }); -->
<!-- </script> -->


\newcommand{\bm}[1]{\boldsymbol{#1}}
\newcommand{\tx}[1]{\textrm{#1}}
\newcommand{\rv}[3]{#1_{#2},\ldots,#1_{#3}}
\newcommand{\RR}{\bm{R}}
\newcommand{\BB}{\bm{B}}
\newcommand{\JJ}{\bm{J}}
\newcommand{\XX}{\bm{X}}
\newcommand{\UU}{\bm{U}}
\newcommand{\VV}{\bm{V}}
\newcommand{\dX}{\Delta\XX}
\newcommand{\YY}{\bm{Y}}
\newcommand{\ZZ}{\bm{Z}}
\newcommand{\dt}{\Delta t}
\newcommand{\tmin}{t_{\tx{min}}}
\newcommand{\tmax}{t_{\tx{max}}}
\newcommand{\aD}{(\alpha,D)}
\newcommand{\eps}{\varepsilon}
\newcommand{\N}{\mathcal{N}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\msd}{\operatorname{MSD}}
\newcommand{\acf}{\operatorname{ACF}}
\newcommand{\var}{\operatorname{var}}
\newcommand{\se}{\operatorname{se}}
\newcommand{\cov}{\operatorname{cov}}
\newcommand{\tr}{\operatorname{trace}}
\newcommand{\arma}{\operatorname{ARMA}}
\newcommand{\snr}{\mathrm{SNR}}
\newcommand{\kbt}{k_{\mathrm{B}}T}
\newcommand{\argmax}{\operatorname{arg\,max}}
\newcommand{\mmu}{\bm{\mu}}
\newcommand{\eet}{\bm{\eta}}
\newcommand{\tth}{\bm{\theta}}
\newcommand{\lla}{\bm{\lambda}}
\newcommand{\oom}{\bm{\omega}}
\newcommand{\pps}{\bm{\psi}}
\newcommand{\pph}{\bm{\phi}}
\newcommand{\zze}{\bm{\zeta}}
\newcommand{\SSi}{\bm{\Sigma}}
\newcommand{\ndim}{d}
\newcommand{\ndrift}{k}
\newcommand*\ud{\mathop{}\!\mathrm{d}}

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# silently load packages not used in code
## suppressMessages({
##   require(kableExtra)
## })
# format link to CRAN packages
cran_link <- function(pkg) {
  paste0("[**", pkg, "**](https://CRAN.R-project.org/package=", pkg, ")")
}
```

# Introduction

The **subdiff** package provides a number of utilities for analyzing subdiffusion in passive particle tracking data.  Let $\XX(t) = \big(X_1(t), \ldots, X_{\ndim}(t)\big)$ denote the $\ndim$-dimensional position of a particle at time $t$, where $\ndim$ is typically between 1 and 3.  In an ideal experiment, it is typically assumed that $\XX(t)$ is a Continuous Stationary Increments (CSI) process, i.e., a stochastic process with mean zero and such that the distribution of $\dX(t) = \XX(t+\dt) - \XX(t)$ is independent of $t$.  Under these conditions, the covariance properties of $\XX(t)$ can often be deduced entirely from the particle's mean square displacement (MSD),
\begin{equation}
\msd_{\XX}(t) = \sum_{i=1}^\ndim E\big[\big(X_i(t)-X_i(0)\big)^2\big].
(\#eq:msd)
\end{equation}
A widely observed phenomenon is that the MSD of microparticles diffusing in biological fluids has a power-law signature on a given timescale,
\begin{equation}
  \msd_{\XX}(t) \sim 2\ndim D \cdot t^\alpha, \qquad t \in (\tmin, \tmax).
  (\#eq:subdiff)
\end{equation}
Whereas $\alpha = 1$ corresponds to the MSD of ordinary diffusion observed in Brownian particles, the power law MSD \@ref(eq:subdiff) with $\alpha \in (0,1)$ is referred to as *subdiffusion*.  

The purpose of the **subdiff** library is to provide simple and computationally efficient tools for modeling and estimating subdiffusion.  An overview of these tools is presented below.  

## Installation

```{r load_packages, include = FALSE, message = FALSE}
# load required packages
require(subdiff)
require(tidyr)
require(dplyr)
require(optimCheck)
require(ggplot2)
require(gridExtra)
## require(RColorBrewer)
## require(parallel)
```

To install the **subdiff** package:

1.  Install the `r cran_link("Rcpp")` package.  To do this properly you will also need to install a C++ compiler.  Instructions for this are available [here](https://teuder.github.io/rcpp4everyone_en/020_install.html).  

    To test that this step is done correctly, quit + restart R, then run the following command:
	
    ```{r rcpp_test, eval = FALSE}
    Rcpp::evalCpp("2 + 2")
    ```

    If the output is `4` then **Rcpp** is installed correctly.

2.  Install the `r cran_link("devtools")` package.

3.  In an R session run the following commands:

    ```{r install, eval =FALSE}
    devtools::install_github("mlysy/subdiff",
                             force = TRUE, INSTALL_opts = "--install-tests")
    ```

4.  Once the packages are installed, to test that everything works correctly, quit + restart R then run the command:

    ```{r subdiff_test, eval = FALSE}
    testthat::test_package("subdiff", reporter = "progress")
    ```
	
	Occasionally due to random chance a few of the tests may fail.  However, if everything works correctly then rerunning this a few times will eventually produce zero test failures.  


```{r hbe_info, include = FALSE}
npaths <- length(unique(hbe$id)) # number of trajectories
n_rng <- range(tapply(hbe$id, hbe$id, length)) # range of trajectory lengths
dt <- 1/60 # interobservation time
```

```{r dataset, child = "_subdiff/_dataset.Rmd"}
```

```{r semiparam, child = "_subdiff/_semiparametric.Rmd"}
```

```{r param, child = "_subdiff/_parametric.Rmd"}
```

```{r csiclass, child = "_subdiff/_csiclass.Rmd"}
```

# Appendix: Code for Figure \@ref(fig:fsdvsfma)

```{r fsdvsfmacode, ref.label = "fsdvsfma", eval = FALSE}
```

# References

